---
title: "Cancer_survival_extrapolation_test_code"
author: "Danielle N"
date: '2022-05-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This code is the test code to develop once data access has been approved. The objectives are:

  * Create a table 1 with descriptives
  * Calculate the overall survival for cancers
  * Using various extrapolation methods test and predict the survival 1,5, 10 years from diagnosis

## Install packages

```{r install_packages,  message=FALSE, warning=FALSE}

rm(list = ls()) # clear environment

library(here)
library(readr)
library(DBI)
library(RSQLite)
library(dplyr)
library(stringr)
library(DT)
library(kableExtra)
library(CodelistGenerator)
library(Eunomia)
library(dbplyr)
library(tictoc)
library(ggplot2)
library(compareGroups)
library(table1)
library(RPostgreSQL)
library(survival)
library(flexsurv)
library(muhaz)
library(survminer)
library(utile.visuals)
library(grid)
library(tidyverse)
library(ggsci)
library(plotly)


```

# plotting survival with different extrapolations functions using the lung dataset
```{r table1,  message=FALSE, warning=FALSE,echo=FALSE}


data(cancer, package="survival")




################################################

#calculate the survival functions and extract the data we need for each method. For the extrapolation functions we want to extrapolate to 10 years

# survival data

t <- seq(0, 3652.5, by=1) # tells the extrapolation to go out to 10 years

# for each function use flexsurvreg to model and then extract the output

Gompertz<-flexsurvreg(Surv(time, status)~1, data=lung, dist="gompertz")
gom_data <-summary(Gompertz,t=t)[[1]] # extract the data
colnames(gom_data) <- paste(colnames(gom_data),"gompertz",sep="_") # add section on column to make unique

weibull <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="weibull")
weib <- summary(weibull,t=t)[[1]]
colnames(weib) <- paste(colnames(weib),"weibull",sep="_") # add section on column to make unique

expo <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="exp")
expo1 <- summary(expo,t=t)[[1]]
colnames(expo1) <- paste(colnames(expo1),"exp",sep="_") # add section on column to make unique

loglog <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="llogis")
loglog1 <- summary(loglog,t=t)[[1]]
colnames(loglog1) <- paste(colnames(loglog1),"loglog",sep="_") # add section on column to make unique

lognorm <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="lnorm")
lognorm1 <- summary(lognorm,t=t)[[1]]
colnames(lognorm1) <- paste(colnames(lognorm1),"lognorm",sep="_") # add section on column to make unique

#spline with 1 knot
one_k_hazard_spline <-
flexsurvspline(formula=Surv(time,status-1)~1,data=lung,k = 1, scale = "hazard")
one_k_hazard_spline1 <- summary(one_k_hazard_spline,t=t)[[1]]
colnames(one_k_hazard_spline1) <- paste(colnames(one_k_hazard_spline1),"Spline1k",sep="_") # add section on column to make unique


#spline with 3 knots
three_k_hazard_spline <- flexsurvspline(formula=Surv(time,status-1)~1,data=lung,k = 3, scale = "hazard")
three_k_hazard_spline1 <- summary(three_k_hazard_spline,t=t)[[1]]
colnames(three_k_hazard_spline1) <- paste(colnames(three_k_hazard_spline1),"Spline3k",sep="_") # add section on column to make unique

gengamma<-flexsurvreg(Surv(time, status)~1, data=lung, dist="gengamma")
gengamma1 <-summary(gengamma,t=t)[[1]] # extract the data
colnames(gengamma1) <- paste(colnames(gengamma1),"gengamma1",sep="_") # add section on column to make unique

# observed
kmsurvival <- survfit (Surv(time, status) ~ 1, data=lung)
km_data <- cbind(kmsurvival$time, kmsurvival$surv, kmsurvival$lower, kmsurvival$upper)
colnames(km_data) <- c("time", "est_km", "lcl_km", "ucl_km")

#combine the extrapolated results together
combined_data <- cbind(weib, gom_data, expo1, loglog1, lognorm1,one_k_hazard_spline1,three_k_hazard_spline1,gengamma1
                       )

#merge with observed
combined_data <- merge(km_data, combined_data, by.x = "time" , by.y = "time_weibull", all = T)
                       
             
# might need to convert days to years
combined_data <- combined_data %>% 
 mutate(Years = round(time_exp/365.25, digit=5))



# pretty plot with all functions together (very messy)
plot_km <- ggplot(combined_data, aes(x = Years, y = est_km)) + 
  xlab("Years") + ylab("Survival Probability") +
  geom_line(aes(y = est_weibull, color="Weibull"), alpha=0.75, size = 1) +
  geom_line(aes(y = est_gompertz , color="Gompertz"), alpha=0.75 , size = 1) +
  geom_line(aes(y = est_exp, color="Exponential" ), alpha=0.75, size = 1) +
  geom_line(aes(y = est_loglog, color="Log-logistic"), alpha=0.75 , size = 1) +
  geom_line(aes(y = est_lognorm, color="Log-normal"), alpha=0.75 , size = 1) +
  geom_line(aes(y = est_Spline1k, color="Spline1k"), alpha=0.75 , size = 1) +
  geom_line(aes(y = est_Spline3k, color="Spline3k"), alpha=0.75 , size = 1) +  
  geom_line(aes(y = est_gengamma1, color="gengamma1"), alpha=0.75 , size = 1) + 
  #geom_ribbon(aes(ymin = lcl_gengamma1, ymax = ucl_gengamma1), alpha=0.1, color="yellow", linetype = 2) + 
  geom_line(data = na.omit(combined_data),aes(y = est_km, color="Observed"), size = 1) +  
  geom_ribbon(data = na.omit(combined_data), aes(ymin = lcl_km, ymax = ucl_km), alpha=0.1, color="grey", linetype = 2) +
  theme_classic() + 
  theme(legend.position = 'top', legend.direction = "horizontal") +
  geom_segment(mapping = aes(x = c(1), y = c(0.5), xend = c(1), yend = c(0)), color = "blue", size = 1 ) +
  geom_segment(mapping = aes(x = c(1), y = c(0.5), xend = c(0), yend = c(0.5)), color = "blue", arrow = arrow(length = unit(0.3, "cm")), size = 1 ) +
  scale_color_identity(guide = "legend") +
  scale_color_manual(name='Extrapolation Model',
                 breaks=c('Observed','Weibull','Gompertz','Exponential','Log-logistic','Log-normal','Spline1k', 'Spline3k', 'gengamma1'), values=c('Observed'='black','Weibull'='blue','Gompertz' = 'pink' ,'Exponential' = 'red','Log-logistic' ='orange', 'Log-normal' ='green', 'Spline1k' = 'purple', 'Spline3k' = 'steelblue', 'gengamma1' = 'yellow' )) +
  geom_text(x=8, y=0.9, label=paste0("Mean OS = ", round(mean(combined_data$est_km,na.rm = T),2) ,
                                       "\nMedian OS = ", round(median(combined_data$est_km, na.rm = T),2)), hjust = 0 ) +
  scale_x_continuous(limits = c(0,max(combined_data$Years)), expand =c(0,0) ,
                     breaks = seq(0,max(combined_data$Years), by = 2 ) ) +
  scale_y_continuous(limits = c(0,1.02), expand =c(0.01,0)) 

plot_km

#ggplotly(plot_km)

############################

# Observed and one function e.g. weibull

plot_km1 <- ggplot(combined_data, aes(x = Years, y = est_km)) + 
  xlab("Years") + ylab("Survival Probability") +
  geom_line(aes(y = est_weibull, color="Weibull"), size = 0.75) +
  geom_ribbon(aes(ymin = lcl_weibull, ymax = ucl_weibull), alpha=0.1, color="steelblue", fill = "steelblue", linetype = 2) +
  geom_line(data = na.omit(combined_data),aes(y = est_km, color="Observed"), size = 0.75) +  
  geom_ribbon(data = na.omit(combined_data), aes(ymin = lcl_km, ymax = ucl_km), alpha=0.1, color="grey50", linetype = 2) +
 theme_bw() + 
  theme(
    
    legend.position = 'top', legend.direction = "horizontal") +
  scale_color_identity(guide = "legend") +
  scale_color_manual(name='Extrapolation Model',
                 breaks=c('Observed','Weibull'), values=c('Observed'='grey50','Weibull'='steelblue')) +
  geom_text(x=7.6, y=0.9, label=paste0("Mean OS = ", round(mean(combined_data$est_km,na.rm = T),2) ,
                                       "\nMedian OS = ", round(median(combined_data$est_km, na.rm = T),2)), hjust = 0 ) +
  scale_x_continuous(limits = c(0,max(combined_data$Years)), expand =c(0,0) ,
                     breaks = seq(0,max(combined_data$Years), by = 2 ) ) +
  scale_y_continuous(limits = c(0,1.02), expand =c(0.01,0)) 


plot_km1

# 
# ###############################
# ## estimating x year survival
# # 1 year
# summary(survfit(Surv(time, status) ~ 1, data = lung), times = 365.25) # can use this to determine where the arrows go on plt
# 
# predict(Gompertz, type = "survival", times = 365.25) %>%
#   slice(1) %>% collect()
# 
# 
# # 5 year
# #summary(kmsurvival, times = 1826.25) # wont work as we dont have data to 5 years
# 
# predict(Gompertz, type = "survival", times = 1826.25) %>%
#   slice(1) %>% collect() # results is naff all people survive
# 
# 
# # 10 year
# #summary(kmsurvival, times = 3652.5)
# 
# predict(Gompertz, type = "survival", times = 3652.5) %>%
#   slice(1) %>% collect() # results is naff all people survive
# 
# 
# 
# 
# #predict mean survival
# predict(Gompertz) %>%
#   slice(1) %>% collect()
# # predict the median survival
# predict(Gompertz, type = "quantile", p = 0.5) %>%
#   slice(1) %>% collect()

```

# plotting cumulative hazard with different extrapolations functions using the lung dataset
```{r cumhaz,  message=FALSE, warning=FALSE,echo=FALSE}

# log cumulative hazard plot

data(lung)

t <- seq(0, 3652.5, by=1) # tells the extrapolation to go out to 10 years

#####################
# log cumulative hazard plot
# 
# extract the cumulative hazard
summary(Gompertz, type = "cumhaz")[[1]]



Gompertz<-flexsurvreg(Surv(time, status)~1, data=lung, dist="gompertz")
gom_data <-summary(Gompertz,t=t , type = "cumhaz")[[1]] # extract the data
colnames(gom_data) <- paste(colnames(gom_data),"gompertz",sep="_") # add section on column to make unique

weibull <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="weibull")
weib <- summary(weibull,t=t, type = "cumhaz")[[1]]
colnames(weib) <- paste(colnames(weib),"weibull",sep="_") # add section on column to make unique

expo <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="exp")
expo1 <- summary(expo,t=t, type = "cumhaz")[[1]]
colnames(expo1) <- paste(colnames(expo1),"exp",sep="_") # add section on column to make unique

loglog <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="llogis")
loglog1 <- summary(loglog,t=t, type = "cumhaz")[[1]]
colnames(loglog1) <- paste(colnames(loglog1),"loglog",sep="_") # add section on column to make unique

lognorm <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="lnorm")
lognorm1 <- summary(lognorm,t=t, type = "cumhaz")[[1]]
colnames(lognorm1) <- paste(colnames(lognorm1),"lognorm",sep="_") # add section on column to make unique

#spline with 1 knot
one_k_hazard_spline <-
flexsurvspline(formula=Surv(time,status-1)~1,data=lung,k = 1, scale = "hazard")
one_k_hazard_spline1 <- summary(one_k_hazard_spline,t=t, type = "cumhaz")[[1]]
colnames(one_k_hazard_spline1) <- paste(colnames(one_k_hazard_spline1),"onekspline",sep="_") # add section on column to make unique

#spline with 3 knots
three_k_hazard_spline <-
flexsurvspline(formula=Surv(time,status-1)~1,data=lung,k = 1, scale = "hazard")
three_k_hazard_spline1 <- summary(three_k_hazard_spline,t=t, type = "cumhaz")[[1]]
colnames(three_k_hazard_spline1) <- paste(colnames(three_k_hazard_spline1),"threekspline",sep="_") # add section on column to make unique



# observed NEED TO WORK OUT HOW TO EXTRACT ci FOR OBSERVED
kmsurvival <- survfit(Surv(time, status) ~ 1, data=lung)
km_data <- cbind(kmsurvival$time, kmsurvival$surv, kmsurvival$lower, kmsurvival$upper)
colnames(km_data) <- c("time", "est_km", "lcl_km", "ucl_km")

#combine the extrapolated results together
combined_data <- cbind(weib, gom_data, expo1, loglog1, lognorm1)

#merge with observed
combined_data <- merge(km_data, combined_data, by.x = "time" , by.y = "time_weibull", all = T)
                       
             
# might need to convert days to years
combined_data <- combined_data %>% 
 mutate(Years = round(time_exp/365.25, digit=5))





```


```{r gof}

###########################################
# extracting stats
#extracts the AIC, BIC, log likihood etc

gof_stats <- rbind(glance(Gompertz),
glance(weibull),
glance(expo) ,
glance(loglog) , 
glance(lognorm),
glance(gengamma) ,
glance(one_k_hazard_spline),
glance(three_k_hazard_spline)

)

# add in column with method name
gof_stats$Method <- c(
  "Gompertz",
  "Weibull",
  "Exponential",
  "Log-Logistic",
  "Log-normal",
  "Generalized Gamma" ,
  "Spline (1 knot)" ,
  "Spline (3 knots)"
)

gof_stats[,c(6:8)] <- round(gof_stats[,c(6:8)],2)

datatable(gof_stats,
          rownames = FALSE ,
          extensions = 'Buttons',
          options = list(
            fixedColumns = TRUE,
            autoWidth = TRUE,
            ordering = FALSE,
            dom = 'tB',
            buttons = c('copy', 'csv', 'pdf')
          
          ) )


```

# plotting function based on multiple extrapolation functions

```{r plotfunction,  message=FALSE, warning=FALSE,echo=FALSE}

# need data in a list format with each list containing the output from a extrapolation object - the original format will be a person inputting the data and then stating which extrapolation functions they want to look at

# INPUT data 
# what survival function(s) do one that says all 
# years of extrapolation

# OUTPUTS
# plots observed versus survival function
# summary AIC/BIC
# cum hazard plots
# other plots


data <- lung
time <- "time"
status <- "status"
extrapolations <- c("gompertz", "weibull", "exp", "llogis", "lnorm", "gengamma")
extrapolations_formatted <- c("Gompertz", "Weibull", "Exponential", "Log-logistic", "Log-normal", "Generalised Gamma")

t <- seq(0, 3652.5, by=1) # tells the extrapolation to go out to 10 years

# for each function use flexsurvreg to model and then extract the output
list_extrap_results <- list()               # Create empty list

for(i in 1:length(extrapolations)) {   # Head of for-loop
  
  model<-flexsurvreg(Surv(time, status)~1, data=data, dist=extrapolations[i])
  model_out <-summary(model,t=t)[[1]] # extract the data
  model_out$Method <- extrapolations_formatted[i]
  list_extrap_results[[i]] <- model_out   # Store output in list
  print(paste0(extrapolations_formatted[i]," ", Sys.time(), " completed"))
}

# get observed data plot
kmsurvival <- survfit (Surv(time, status) ~ 1, data=lung)
km_data <- as.data.frame(cbind(kmsurvival$time, kmsurvival$surv, kmsurvival$lower, kmsurvival$upper))
colnames(km_data) <- c("time", "est", "lcl", "ucl")
km_data$Method <- "Observed"

#get some colours
cols <- c("#00468BFF", #dark blue
         "#ED0000FF", # red
         "#42B540FF", #green
         "#0099B4FF", #lightblue
         "#925E9FFF", # purple
         "#FF6F0EFF", #orange
         "#E377C2FF", #pink
         "#BCBD22FF", #olive
         "#AD002AFF") # dark red


# plot

for(i in 1:length(extrapolations)) { 

my_colors <- c("darkgrey", cols[i])

# for each extrapolation method rbind with each observed and create a plot
extrap_results <- list_extrap_results[[i]]

#rbind the observed results to the extrapolated ones
extrap_results1 <- rbind(km_data, extrap_results)

extrap_results1$Method <- factor(extrap_results1$Method, levels=c('Observed', extrap_results1$Method[nrow(extrap_results1)] ))
             
# might need to convert days to years
extrap_results1 <- extrap_results1 %>% 
 mutate(Years = round(time/365.25, digit=5))

plot_km1 <- ggplot(extrap_results1, aes(x = Years, y = est, colour = Method)) + 
  xlab("Years") + ylab("Survival Probability") +
  geom_line() +
  geom_ribbon(aes(ymin = lcl, ymax = ucl, fill = Method), linetype = 2, alpha = 0.1) +
  scale_color_manual(values = my_colors) +
  scale_fill_manual(values = my_colors) +
  theme_bw() + 
  theme( legend.position = 'top', legend.direction = "horizontal") +
  scale_x_continuous(limits = c(0,max(combined_data$Years)), expand =c(0,0) ,
                     breaks = seq(0,max(combined_data$Years), by = 2 ) ) +
  scale_y_continuous(limits = c(0,1.02), expand =c(0.01,0)) 

ggsave(plot_km1, file= here(paste0("plot_survival", extrapolations_formatted[i],".png"))
          , width = 14, height = 10, units = "cm")

}

fig <- ggplotly(plot_km1)

fig <- plotly_build(plot_km1)
str(fig)
 
 
 fig$x$data[[1]]$name <- 'Observed'
 fig$x$data[[2]]$name <- 'Exponential'
fig$x$data[[3]]$name <- NA
 fig$x$data[[4]]$name <- NA
 
fig$x$data[[1]]$legendgroup
fig$x$data[[2]]$legendgroup
fig$x$data[[3]]$legendgroup <- NA
fig$x$data[[4]]$legendgroup <- NA



  geom_text(x=7.6, y=0.9, label=paste0("Mean OS = ", round(mean(combined_data$est_km,na.rm = T),2) ,
                                       "\nMedian OS = ", round(median(combined_data$est_km, na.rm = T),2)), hjust = 0 ) +



ggplotly(plot_km) # keeps the CI in the same place with no extra button

```

```{r interactiveplot,  message=FALSE, warning=FALSE,echo=FALSE}

data(lung)

################################################

#calculate the survival functions and extract the data we need for each method. For the extrapolation functions we want to extrapolate to 10 years

# survival data

t <- seq(0, 3652.5, by=1) # tells the extrapolation to go out to 10 years

# for each function use flexsurvreg to model and then extract the output

Gompertz<-flexsurvreg(Surv(time, status)~1, data=lung, dist="gompertz")
gom_data <-summary(Gompertz,t=t)[[1]] # extract the data
colnames(gom_data) <- paste(colnames(gom_data),"gompertz",sep="_") # add section on column to make unique

weibull <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="weibull")
weib <- summary(weibull,t=t)[[1]]
colnames(weib) <- paste(colnames(weib),"weibull",sep="_") # add section on column to make unique

expo <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="exp")
expo1 <- summary(expo,t=t)[[1]]
colnames(expo1) <- paste(colnames(expo1),"exp",sep="_") # add section on column to make unique

loglog <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="llogis")
loglog1 <- summary(loglog,t=t)[[1]]
colnames(loglog1) <- paste(colnames(loglog1),"loglog",sep="_") # add section on column to make unique

lognorm <- flexsurvreg(Surv(time,status-1)~1,data=lung,dist="lnorm")
lognorm1 <- summary(lognorm,t=t)[[1]]
colnames(lognorm1) <- paste(colnames(lognorm1),"lognorm",sep="_") # add section on column to make unique

#spline with 1 knot
one_k_hazard_spline <-
flexsurvspline(formula=Surv(time,status-1)~1,data=lung,k = 1, scale = "hazard")
one_k_hazard_spline1 <- summary(one_k_hazard_spline,t=t)[[1]]
colnames(one_k_hazard_spline1) <- paste(colnames(one_k_hazard_spline1),"Spline1k",sep="_") # add section on column to make unique


#spline with 3 knots
three_k_hazard_spline <- flexsurvspline(formula=Surv(time,status-1)~1,data=lung,k = 3, scale = "hazard")
three_k_hazard_spline1 <- summary(three_k_hazard_spline,t=t)[[1]]
colnames(three_k_hazard_spline1) <- paste(colnames(three_k_hazard_spline1),"Spline3k",sep="_") # add section on column to make unique

gengamma<-flexsurvreg(Surv(time, status)~1, data=lung, dist="gengamma")
gengamma1 <-summary(gengamma,t=t)[[1]] # extract the data
colnames(gengamma1) <- paste(colnames(gengamma1),"gengamma1",sep="_") # add section on column to make unique

# observed
kmsurvival <- survfit (Surv(time, status) ~ 1, data=lung)
km_data <- cbind(kmsurvival$time, kmsurvival$surv, kmsurvival$lower, kmsurvival$upper)
colnames(km_data) <- c("time", "est_km", "lcl_km", "ucl_km")

#combine the extrapolated results together
combined_data <- cbind(weib, gom_data, expo1, loglog1, lognorm1,one_k_hazard_spline1,three_k_hazard_spline1,gengamma1
                       )

#merge with observed
combined_data <- merge(km_data, combined_data, by.x = "time" , by.y = "time_weibull", all = T)
                       
             
# might need to convert days to years
combined_data <- combined_data %>% 
 mutate(Years = round(time_exp/365.25, digit=5))

# interactive figure
fig <- plot_ly(combined_data, x = ~Years, y = ~est_km, name = 'Observed', type = 'scatter', mode = 'lines', connectgaps = TRUE) 

fig <- fig %>% add_trace(combined_data, x = ~Years, y = ~ucl_km, type = 'scatter', mode = 'lines',
        line = list(color = 'rgba(0,100,80,1)'),
        showlegend = FALSE, name = 'High KM')
fig <- fig %>% add_trace(combined_data, x = ~Years, y = ~lcl_km, type = 'scatter', mode = 'lines',
        line = list(color = 'rgba(0,100,80,1)'),
        showlegend = FALSE, name = 'low KM')
fig <- fig %>% add_trace(y = ~est_weibull, name = 'Weibull', mode = 'lines') 
fig <- fig %>% add_trace(y = ~est_gompertz, name = 'Gompertz', mode = 'lines') 
fig <- fig %>% add_trace(y = ~est_exp, name = 'Exponential', mode = 'lines') 
fig <- fig %>% add_trace(y = ~est_loglog, name = 'Log-Logistic', mode = 'lines') 
fig <- fig %>% add_trace(y = ~est_lognorm, name = 'Log-Normal', mode = 'lines') 
fig <- fig %>% add_trace(y = ~est_gengamma1, name = 'Generalised Gamma', mode = 'lines') 
fig <- fig %>% add_trace(y = ~est_Spline1k, name = 'Spline (1 Knot)', mode = 'lines') 
fig <- fig %>% add_trace(y = ~est_Spline3k, name = 'Spline (3 Knots)', mode = 'lines') 
fig <- fig %>% layout(xaxis = list(title = "Years"),
         yaxis = list (title = "Survival Probability"))

fig


# pretty plot with all functions together (very messy)
plot_km <- ggplot(combined_data, aes(x = Years, y = est_km)) + 
  xlab("Years") + ylab("Survival Probability") +
  geom_line(aes(y = est_weibull, color="Weibull"), alpha=0.75, size = 1) +
  geom_line(aes(y = est_gompertz , color="Gompertz"), alpha=0.75 , size = 1) +
  geom_line(aes(y = est_exp, color="Exponential" ), alpha=0.75, size = 1) +
  geom_line(aes(y = est_loglog, color="Log-logistic"), alpha=0.75 , size = 1) +
  geom_line(aes(y = est_lognorm, color="Log-normal"), alpha=0.75 , size = 1) +
  geom_line(aes(y = est_Spline1k, color="Spline1k"), alpha=0.75 , size = 1) +
  geom_line(aes(y = est_Spline3k, color="Spline3k"), alpha=0.75 , size = 1) +  
  geom_line(aes(y = est_gengamma1, color="gengamma1"), alpha=0.75 , size = 1) + 
  geom_line(data = na.omit(combined_data),aes(y = est_km, color="Observed"), size = 1) +  
  geom_ribbon(data = na.omit(combined_data), aes(ymin = lcl_km, ymax = ucl_km), alpha=0.1, color="grey", linetype = 2) +
  theme_classic() + 
  theme(legend.position = 'top', legend.direction = "horizontal") +
  geom_segment(mapping = aes(x = c(1), y = c(0.5), xend = c(1), yend = c(0)), color = "blue", size = 1 ) +
  geom_segment(mapping = aes(x = c(1), y = c(0.5), xend = c(0), yend = c(0.5)), color = "blue", arrow = arrow(length = unit(0.3, "cm")), size = 1 ) +
  scale_color_identity(guide = "legend") +
  scale_color_manual(name='Extrapolation Model',
                 breaks=c('Observed','Weibull','Gompertz','Exponential','Log-logistic','Log-normal','Spline1k', 'Spline3k', 'gengamma1'), values=c('Observed'='black','Weibull'='blue','Gompertz' = 'pink' ,'Exponential' = 'red','Log-logistic' ='orange', 'Log-normal' ='green', 'Spline1k' = 'purple', 'Spline3k' = 'steelblue', 'gengamma1' = 'yellow' )) +
  geom_text(x=8, y=0.9, label=paste0("Mean OS = ", round(mean(combined_data$est_km,na.rm = T),2) ,
                                       "\nMedian OS = ", round(median(combined_data$est_km, na.rm = T),2)), hjust = 0 ) +
  scale_x_continuous(limits = c(0,max(combined_data$Years)), expand =c(0,0) ,
                     breaks = seq(0,max(combined_data$Years), by = 2 ) ) +
  scale_y_continuous(limits = c(0,1.02), expand =c(0.01,0)) 

plot_km



```{r sessioninfor}

# print out the session information
sessionInfo()

#dbDisconnect(db)

```
